---
title: "GettingStarted"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GettingStarted}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Installing and using the package.  
# The following three lines only need to be ran only once!

```{r eval=FALSE}
install.packages("devtools")
library("devtools")

devtools::install_github("javirudolph/StabilityToolkit")
```

Now that the package is installed in your computer, you can load it to your R session:
```{r}
library(StabilityToolkit)
```

```{r eval=FALSE}
# Check the help file
help(package="StabilityToolkit")
```

For this short example, we will use the `G1000` data set, which is included in the package.
```{r}
# You can load the data:
data("G1000")

# We will call the data `trial.data` for this example
trial.data <-G1000
print(trial.data)
```

 Now fitting the segregation-selection (SS) model to this trial data set
```{r}
simpleSSfit <- dynamic.fit(data1=trial.data,model="SS")
```


Try fitting the HT model
```{r}
simpleHTfit <- dynamic.fit(data1=trial.data,model="HT")
```


Adding complexity: 
1. Assume the cost parameter is negative: i.e., bacteria with plasmid grows faster.  Just add
```{r}
simpleSSfit <- dynamic.fit(data1=trial.data,model="SS", true.cost=FALSE)
```

2. Assume you have two stability assay data sets and want to test if they are driven by the same biological parameters.  We can do that for the SS and for the HT models

```{r}
trial.data2 <- trial.data # just use the same data as your second data set, for the sake of this example

comb.SSfit <- dynamic.fit(data1=trial.data,model="SS", comb=TRUE,data2=trial.data2)

comb.HTfit <- dynamic.fit(data1=trial.data,model="HT", comb=TRUE,data2=trial.data2)
```


### We can also do this using two other data sets included in the package:

```{r}
trial.data <- data("L1I1HancPac")
trial.data2 <- data("L1I1HancPev")
```

```{r}
comb.SSfit <- dynamic.fit(data1=trial.data,model="SS", comb=TRUE,data2=trial.data2)

comb.HTfit <- dynamic.fit(data1=trial.data,model="HT", comb=TRUE,data2=trial.data2)
```



3. Assume you have data from a competition assay and from a stability assay and you want to leverage the information in the competition assay to better estimate the cost.  This package comes with two sample data sets to that effect, one is a stability assay, called data.compet and another called data.stabexp. These two data sets come from an Ancestral line, so cost is expected to be true cost. Check them out by typing:
```{r}
data("data.compet")
data.compet

data("data.stabexp")
data.stabexp
```

```{r}
add.compet.fit <- dynamic.fit(data1=data.stabexp,model="SS",add.compet.data=TRUE, data2=data.compet)
```


