---
title: "GettingStarted"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GettingStarted}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Preamble

If you have questions and run into problems, please e-mail me at josemi@ufl.edu. Mathematics and statistics are fantastic tools to translate fundamental questions in biology into testable predictions. It's the best mechanism that we have in science to formulate relevant biological
questions in a way that these can be directly confronted with the data sets at hand. If you have multiple tentative explanations for your results and you speak even just a little of the language of math, you have what you need to translate those tentative explanations into
statistical hypotheses that can all be simultaneously tested against your data. So if you want to make the most of this package, you should familiarize yourself with the models in the [Top lab](http://www.thetoplab.org/). A lot of time and effort went into writing the models in a way that they are easily understood. A great place to start is the paper by [De Gelder et al 2004](https://www.genetics.org/content/168/3/1131) and the model
section of the paper by [Ponciano et al 2007](https://doi.org/10.1534/genetics.106.061937). There you will find all the models used in this
package. Reading those papers is a pre-requisite to use this package smartly, and not just as a plug-and-play tool. Even if you don't fully understand all of the statistical details, you won't regret spending time on these papers. It is my hope that after reading these you will have what you need to de-bug your analyses. 

## Installation

You can install this version from [GitHub](https://github.com/) with:

``` {r eval=FALSE}
# If you have not used the `devtools` package before, you can install it following the `install.packages` function

install.packages("devtools")

devtools::install_github("javirudolph/StabilityToolkit")
``` 

```{r setup}
library(StabilityToolkit)
```

